---
title: "BJ5Ta_RNA_DESeq"
author: "Joanna Yeung"
date: "2025-06-20"
output: html_document
---
```{r}
library("msigdbr")
library("DESeq2")
library("gplots")
library("RColorBrewer")
library("pheatmap")
library("gtools")
library("dplyr")
library("tibble")
library("readxl")
library("enrichR")
library("fgsea")
library("cowplot")
```

```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_1/RNAseq/post_kallisto_analysis/workspaces/genes_anno_type.RData")
```

# annotation colors for pheatmap
```{r heatmap_colors}
hm_cols_RNA <- list(
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "TNFa"="darkblue"), 
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                SASP = c("Yes"="dark green", "No"="white"),
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
```

```{r}
colors <- c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "TNFa"="darkblue") 
```

```{r}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Different_Cell_Lines_NFkB_Profiling_PDvsDoxo_Jan2026/STAR_alignment")
samples <- dir(pattern="ReadsPerGene.out.tab")
samples <- samples[grepl("BJ-5Ta", samples)]
samples <- samples[c(1:3,6:7,4:5,8,9:10)]
samplenames <- gsub("_[0-9]{1,2}_ReadsPerGene.out.tab", "", samples)
geneCounts <- lapply(samples, function(x){df <- read.table(x, skip=4)[ ,c(1,2)]
                                          colnames(df) <- c("geneID", x)
                                          return(df)})
names(geneCounts) <- samplenames

countsMatrix <- data.frame(geneID=geneCounts[[1]]$geneID)
for(i in 1:length(geneCounts)){
  countsMatrix <- merge(countsMatrix, geneCounts[[i]], by="geneID")
}
row.names(countsMatrix) <- countsMatrix$geneID
countsMatrix <- countsMatrix[ ,-1]
colnames(countsMatrix) <- samplenames
metaData <- data.frame(Samples=names(geneCounts), Treatment=c(rep("Cycling",2),rep("Doxo",3), rep("Palbo",3), rep("TNFa",2)), 
                       Biorep=factor(c(1,2,1,2,3,1,2,3,1,2)))

BJ_RNAdds <- DESeqDataSetFromMatrix(countsMatrix, metaData, ~Biorep+Treatment)
BJ_RNAdds <- estimateSizeFactors(BJ_RNAdds)

rlog_BJ_RNAdds <- rlog(BJ_RNAdds)
BJ_rlog_Mat <- assays(rlog_BJ_RNAdds)[[1]]
row.names(BJ_rlog_Mat) <- gsub('\\.[0-9]*$', '', rownames(rlog_BJ_RNAdds))
row.names(BJ_rlog_Mat) <- ifelse(genes_anno_type[match(row.names(BJ_rlog_Mat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(BJ_rlog_Mat), genes_anno_type[match(row.names(BJ_rlog_Mat), genes_anno_type$ensembl_id), ]$gene_id)
# need to save to pdf
plot_grid(
plotPCA(rlog_BJ_RNAdds, intgroup=c("Biorep")), 
plotPCA(rlog_BJ_RNAdds, intgroup=c("Treatment"))
)

# Extract PCA data
pcaData_allSamples <- plotPCA(rlog_BJ_RNAdds, intgroup = c("Treatment"))
percentVar_allSamples <- round(100 * attr(pcaData_allSamples , "percentVar"))
```

# DESeq with LRT test
```{r}
# combine experimental design levels to make it easier to extract results for pairwise comparisons of samples
keep <- rowSums(counts(BJ_RNAdds, normalized=T)) >= 1 # filter out rows with rowsum less than 1 to remove genes that aren't being expressed across samples. 
BJ_RNAdds <- BJ_RNAdds[keep,]
BJ_RNAddsallLRT <- DESeq(BJ_RNAdds, test="LRT", reduced=~Biorep) #perform DESeq
#extract results from DESeq objects 

# get DESeq results from LRT
acrossBJ.Groups <- results(BJ_RNAddsallLRT)
acrossBJ.Groups$SYMBOL <-  gsub('\\.[0-9]*$', '', row.names(acrossBJ.Groups))
acrossBJ.Groups$SYMBOL <-  ifelse(genes_anno_type[match(acrossBJ.Groups$SYMBOL, genes_anno_type$ensembl_id), ]$gene_id %in% NA, acrossBJ.Groups$SYMBOL, genes_anno_type[match(acrossBJ.Groups$SYMBOL, genes_anno_type$ensembl_id), ]$gene_id)

# filter for significant results (padj < 0.05)
sigBJ.Groups <- acrossBJ.Groups[acrossBJ.Groups$padj < 0.05 & !is.na(acrossBJ.Groups$padj), ]

sigBJ.Groups$SYMBOL <-  ifelse(genes_anno_type[match(sigBJ.Groups$SYMBOL, genes_anno_type$ensembl_id), ]$gene_id %in% NA, sigBJ.Groups$SYMBOL, genes_anno_type[match(sigBJ.Groups$SYMBOL, genes_anno_type$ensembl_id), ]$gene_id)

sigBJ.Groups <- sigBJ.Groups[order(sigBJ.Groups$padj), ]
```

```{r}
BJ_RNAdds <- DESeq(BJ_RNAdds)
LFCResults_BJ_RNA <- list()
# comparison with Cycling as control
for(i in 1:3){
LFCResults_BJ_RNA[[i]] <- results(BJ_RNAdds, contrast=c("Treatment", as.character(levels(BJ_RNAdds$Treatment)[i+1]), as.character(levels(BJ_RNAdds$Treatment)[1])))
                                       
row.names(LFCResults_BJ_RNA[[i]]) <- gsub('\\.[0-9]*$', '', row.names(LFCResults_BJ_RNA[[i]]))

LFCResults_BJ_RNA[[i]]$SYMBOL <-  ifelse(genes_anno_type[match(row.names(LFCResults_BJ_RNA[[i]]), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(LFCResults_BJ_RNA[[i]]), genes_anno_type[match(row.names(LFCResults_BJ_RNA[[i]]), genes_anno_type$ensembl_id), ]$gene_id)

names(LFCResults_BJ_RNA)[[i]] <- paste(as.character(levels(BJ_RNAdds$Treatment)[i+1]), "vs", as.character(levels(BJ_RNAdds$Treatment)[1]))
}
```

# filter for significant results with absolute log2FC > 1 across all pairwise comparisons between groups. 
```{r}
sigLFCResults_BJ_RNA <- list()
for(i in 1:length(LFCResults_BJ_RNA)){
  sigLFCResults_BJ_RNA[[i]] <- LFCResults_BJ_RNA[[i]][LFCResults_BJ_RNA[[i]]$SYMBOL %in% sigBJ.Groups$SYMBOL& abs(LFCResults_BJ_RNA[[i]]$log2FoldChange) >= 1, ]
  sigLFCResults_BJ_RNA[[i]]$LRT_padj <- acrossBJ.Groups[match(row.names(sigLFCResults_BJ_RNA[[i]]), row.names(acrossBJ.Groups)), ]$padj
}
names(sigLFCResults_BJ_RNA) <- names(LFCResults_BJ_RNA)

# get list of genes with absolute log2FC > 1 increase relative to controls by extracting unique genes from sigLFCResults
sig.BJ.Genes <- list()
for(i in 1:length(sigLFCResults_BJ_RNA)){
sig.BJ.Genes[[i]] <- sigLFCResults_BJ_RNA[[i]]$SYMBOL
}
sig.BJ.Genes <- unique(unlist(sig.BJ.Genes))
as.data.frame(colData(BJ_RNAdds))
```

# download msigdb databases
```{r}
# msigdb Hmark database
msigdbr_df_H <- as.data.frame(msigdbr(species = "Homo sapiens", collection="H"))
msigdbr_t2g_H = msigdbr_df_H %>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()
msigdbr_t2g_H$gs_name <- sub("HMARK_", "", msigdbr_t2g_H$gs_name)
msigdbr_t2g_H$gs_name <- gsub("_", " ", msigdbr_t2g_H$gs_name)

# msigdb KEGG pathway database
msigdbr_df_KEGG <- as.data.frame(msigdbr(species = "Homo sapiens", collection="C2", subcollection ="KEGG_MEDICUS"))
msigdbr_t2g_KEGG = msigdbr_df_KEGG %>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()
msigdbr_t2g_KEGG$gs_name <- sub("KEGG_", "", msigdbr_t2g_KEGG$gs_name)
msigdbr_t2g_KEGG$gs_name <- gsub("_", " ", msigdbr_t2g_KEGG$gs_name)

# msigdb Reactome database
msigdbr_df_REACT <- as.data.frame(msigdbr(species = "Homo sapiens", collection="C2", subcollection="CP:REACTOME"))
msigdbr_t2g_REACT = msigdbr_df_REACT %>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()
msigdbr_t2g_REACT$gs_name <- sub("REACTOME_", "", msigdbr_t2g_REACT$gs_name)
msigdbr_t2g_REACT$gs_name <- gsub("_", " ", msigdbr_t2g_REACT$gs_name)
```

# load SASP gene list
```{r}
# SASP genes database downloaded from Wang et al, EMBO 2022. 
Wang_etal_SASP <- as.data.frame(read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_1/RNAseq/post_kallisto_analysis/Other_Resources/embj2021108946-sup-0003-tableev2.xlsx"))
Wang_etal_SASP <- genes_anno_type[match(Wang_etal_SASP$p53, genes_anno_type$gene_id), ] %>% dplyr::distinct(gene_id, ensembl_id, entrezid)
Wang_etal_SASP$Type <- ifelse(Wang_etal_SASP$gene_id %in% unlist(strsplit("BMP2\nLIPG\nIL1A\nAPOE\nC3\nB2M\nCCL2\nCCL5\nCXCL14\nCXCL16\nCXCL2\nCXCL8\nEBI3\nF8\nFas\nIL18\nIL1B\nIL6\nNRG1\nPENK\nSAA1\nTNC\nTNFSF15\nVEGFC\nMMP3\nBDNF\nMMP1\nKISS1\nFN1\nEDN1\nPLAU\nIL11\nCXCL1", "\n")), "NF-KB", "")

Wang_etal_SASP$Type <- ifelse(Wang_etal_SASP$gene_id %in% unlist(strsplit("VGF\nAPOD\nCEACAM1\nCPE\nNTN1\nGDF15\nAGT\nLAMA3\nCLCA2\nPLA2G15\nISG15\nLIF\nMR1\nPLTP\nTSKU\nPLAT\nLYNX1\nTGFB2\nTGFA\nAGRN\nTGFBI\nVWCE\nIGFBP3\nGAS6\nPCSK5\nPLXNB1\nHBEGF\nABHD15\nLAMC1\nMFGE8\nPGF\nWNT9A\nVCAN\nTIMP3", "\n")), "p53", ifelse(Wang_etal_SASP$gene_id %in% unlist(strsplit("BMP2\nLIPG\nIL1A\nAPOE\nC3\nB2M\nCCL2\nCCL5\nCXCL14\nCXCL16\nCXCL2\nCXCL8\nEBI3\nF8\nFas\nIL18\nIL1B\nIL6\nNRG1\nPENK\nSAA1\nTNC\nTNFSF15\nVEGFC\nMMP3\nBDNF\nMMP1\nKISS1\nFN1\nEDN1\nPLAU\nIL11\nCXCL1", "\n")), "NF-KB", ""))

# SASP genes from Gleason et al, Clincal Cancer Research
Gleason_etal_SASP <- as.data.frame(read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/Gleason_etal_ClinicalCancerRes2024_SASPList_supplementary_table_s6.xlsx",col_names = FALSE))

# SASP genes from SASP Atlas
SASP_Atlas_sigMat <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/BJ_Lauren_bulkRNAseq2022_analysis/databases/SASP_Atlas.csv", sep=",", header=T)

# merge all sources of SASP genes into a single vector
SASP_all <- unique(c(as.character(SASP_Atlas_sigMat$Genes), Gleason_etal_SASP$...1, Wang_etal_SASP$gene_id))
```

# load p53 & NF-kB target genes list 
```{r}
# import p53 census target gene database from https://tp53.isb-cgc.org/target_genes
Fischer_Oncogene_2017_p53_census_target_genes <- read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/Fischer_Oncogene_2017_p53_census_target_genes.xlsx")

# import Gilmore Lab's list of NF-kB target genes
Gilmore_NFKB_Targets <- as.data.frame(read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/GilmoreLab_NFKB_Target_Genes_BostonUniversity.xlsx"))$`Human Gene Name`
Gilmore_NFKB_Targets <- unique(Gilmore_NFKB_Targets[!is.na(Gilmore_NFKB_Targets)])

# extract genes under Hallmark terms P53 PATHWAY & TNFA SIGNALING VIA NFKB
p53_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_P53_PATHWAY", ]$gene_symbol
NFKB_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_TNFA_SIGNALING_VIA_NFKB", ]$gene_symbol
```

# annotate genes
```{r}
BJ.anno_genes <- data.frame(row.names=unique(sigBJ.Groups$SYMBOL))

BJ.anno_genes$Cell_Cycle <- ifelse(row.names(BJ.anno_genes) %in% c(msigdbr_df_H[msigdbr_df_H$gs_name %in% c("HALLMARK_E2F_TARGETS", "HALLMARK_G2M_CHECKPOINT"), ]$gene_symbol, msigdbr_df_REACT[msigdbr_df_REACT$gs_name %in% c("REACTOME_CELL_CYCLE", "REACTOME_CELL_CYCLE_CHECKPOINTS", "REACTOME_CELL_CYCLE_MITOTIC"), ]$gene_symbol, msigdbr_df_KEGG[msigdbr_df_KEGG$gs_name %in% "KEGG_CELL_CYCLE", ]$gene_symbol), "Yes", "No")

# add annotation to if genes are SASP
BJ.anno_genes$SASP <- ifelse(row.names(BJ.anno_genes) %in% SASP_all, "Yes", "No")

# add annotation to significant genes based on if they are NF-kB or p53 targets.
BJ.anno_genes$NFKB_All <- ifelse(row.names(BJ.anno_genes) %in% c(Gilmore_NFKB_Targets, NFKB_Hallmark, Wang_etal_SASP[Wang_etal_SASP$Type %in% "NF-KB", ]$gene_id), "Yes", "No") 
BJ.anno_genes$p53_All <- ifelse(row.names(BJ.anno_genes) %in% c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark, Wang_etal_SASP[Wang_etal_SASP$Type %in% "p53", ]$gene_id), "Yes", "No")

# extract rlog counts of sig changing genes
BJ.sigMat_0.05 <- BJ_rlog_Mat[row.names(BJ_rlog_Mat) %in% sig.BJ.Genes, ]
```

# plot heatmap of sig genes in BJ
```{r}
BJ.annoCol <- as.data.frame(colData(BJ_RNAdds))
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Different_Cell_Lines_NFkB_Profiling_PDvsDoxo_Jan2026/post_STAR_alignment_analysis/Figures"
```
```{r}
pdf(paste0(pdf_dir, "/BJ.sigMat_0.05.log2FCover1.pdf"))
bad_rows <- apply(BJ.sigMat_0.05, 1, function(x) {
  any(is.na(x) | is.nan(x) | is.infinite(x)) || sd(x) == 0
})

BJ.sigMat_clean <- BJ.sigMat_0.05[!bad_rows, ]
pheatmap(BJ.sigMat_clean, scale="row", cluster_cols = F, cluster_rows = T, show_rownames = F, annotation_row = BJ.anno_genes, annotation_col=BJ.annoCol[ ,2:3] , annotation_colors = hm_cols_RNA, breaks=seq(-2.5, 2.5, length.out = 100))
dev.off()
# plot NFkB target genes
pdf(paste0(pdf_dir, "/BJ.sigMat_0.05.log2FCover1.NFKBtargets.pdf"), height=14)
pheatmap(BJ.sigMat_0.05[row.names(BJ.sigMat_0.05) %in% row.names(BJ.anno_genes[BJ.anno_genes$NFKB_All %in% "Yes", ]), ], scale = "row",  cluster_cols = F, cluster_rows = T, show_rownames = T, annotation_row = BJ.anno_genes, annotation_col=BJ.annoCol[ ,2:3] , annotation_colors = hm_cols_RNA, breaks=seq(-2.5, 2.5, length.out = 100))
dev.off()
```

# compare NFkB target genes changing in BJs vs LS8817s
```{r}
# load workspace that contains RNA-seq analysis in LS8817s
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure2/workspace.RData")
# add annotation to significant genes based on if they are NF-kB or p53 targets. 
clusterDF$NFKB_All <- ifelse(clusterDF$gene_id %in% c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB), "Yes", "No") # annotate sig genes as NF-kB targets if they are part of Gilmore_NFKB_Targets or TNFA SIGNALING VIA NFKB Hallmark term 
clusterDF$p53_All <- ifelse(clusterDF$gene_id %in% c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark), "Yes", "No")

LS8817_NFkB <- list(Palbo=clusterDF[clusterDF$Clusters %in% c("7", "1") & clusterDF$NFKB_All %in% "Yes", ],
                    Doxo=clusterDF[clusterDF$Clusters %in% c("4", "6") & clusterDF$NFKB_All %in% "Yes", ], 
                    Shared=clusterDF[clusterDF$Clusters %in% "3" & clusterDF$NFKB_All %in% "Yes", ]
)
BJ_NFkB <- lapply(sigLFCResults_BJ_RNA, function(x){x[x$SYMBOL %in% row.names(BJ.anno_genes[BJ.anno_genes$NFKB_All %in% "Yes", ]) & x$log2FoldChange >0, ]$SYMBOL})
```

# RNA-seq: plot venn diagram showing overlap of NF-kB target genes going up in LS8817 vs BJ cells
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2))
}
```

# get list of genes in the venn diagram overlap
```{r}
get_intersections_named <- function(sets) {
  n <- length(sets)
  set_names <- names(sets)
  
  if (is.null(set_names)) {
    stop("Please provide named sets (e.g., list(A=..., B=...))")
  }
  
  all_combinations <- unlist(
    lapply(1:n, function(x) combn(n, x, simplify = FALSE)),
    recursive = FALSE
  )
  
  result <- list()
  
  for (combo in all_combinations) {
    
    combo_names <- set_names[combo]
    name <- paste(combo_names, collapse = "&")
    
    intersect_genes <- Reduce(intersect, sets[combo])
    
    other_sets <- setdiff(seq_len(n), combo)
    if (length(other_sets) > 0) {
      other_genes <- unique(unlist(sets[other_sets]))
      intersect_genes <- setdiff(intersect_genes, other_genes)
    }
    
    # 🔥 skip empty overlaps
    if (length(intersect_genes) == 0) next
    
    df <- data.frame(
      gene_id = intersect_genes,
      intersect_cat = name,
      stringsAsFactors = FALSE
    )
    
    result[[name]] <- df
  }
  
  result_df <- do.call(rbind, result)
  rownames(result_df) <- NULL
  
  return(result_df)
}
```
```{r}
venn_data_shared <- list(LS8817_Doxo=LS8817_NFkB$Doxo$gene_id, BJ_Doxo14=BJ_NFkB$`Doxo vs Cycling`,BJ_TNFa=BJ_NFkB$`TNFa vs Cycling`)
LS8817vsBJ_Doxo.df <- get_intersections_named(venn_data_shared)
LS8817vsBJ_Doxo.df$SASP <- ifelse(LS8817vsBJ_Doxo.df$gene_id %in% row.names(BJ.anno_genes[BJ.anno_genes$SASP %in% "Yes", ]), "1", "0")


pdf(paste0(pdf_dir, "/LS8817vsBJ_Doxo.NFKB.genes_vennDiagram.pdf"), width=4, height=10)
create_proportional_venn(sets=venn_data_shared, 
                         set_labels = c("LS8817 Doxo", "BJ Doxo", "BJ +TNFa"), 
                         colors = c("#B3C543","#D1DC8E",  "darkblue")
)
ggplot(LS8817vsBJ_Doxo.df,
       aes(x = intersect_cat,
           y = gene_id,
           color = SASP,
           label = gene_id)) +
  geom_text(size = 3) +
  facet_wrap(~intersect_cat, scales="free")+
  scale_color_manual(values = c("0" = "black", "1" = "darkgreen")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(), 
    legend.position = "none"
  ) +
  labs(x = "", y = "")
dev.off()
```
```{r}
pdf(paste0(pdf_dir, "/LS8817vsBJ_Palbo_NFKB.genes_vennDiagram.pdf"), width=4, height=10)
venn_data_shared <- list(LS8817_Palbo=LS8817_NFkB$Palbo$gene_id, BJ_Palbo=BJ_NFkB$`Palbo vs Cycling`, BJ_TNFa=BJ_NFkB$`TNFa vs Cycling`)
create_proportional_venn(sets=venn_data_shared, 
                         set_labels = c("LS8817 Palbo", "BJ Palbo", "BJ +TNFa"), 
                         colors = c("#E0ADCD", "#ECCEE1" ,"darkblue") 
)
LS8817vsBJ_Palbo.df <- get_intersections_named(venn_data_shared)
LS8817vsBJ_Palbo.df$SASP <- ifelse(LS8817vsBJ_Palbo.df$gene_id %in% row.names(BJ.anno_genes[BJ.anno_genes$SASP %in% "Yes", ]), "1", "0")

ggplot(LS8817vsBJ_Palbo.df,
       aes(x = intersect_cat,
           y = gene_id,
           color = SASP,
           label = gene_id)) +
  geom_text(size = 3) +
  facet_wrap(~intersect_cat, scales="free")+
  scale_color_manual(values = c("0" = "black", "1" = "darkgreen")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(), 
    legend.position = "none"
  ) +
  labs(x = "", y = "")
dev.off()
```

```{r}
pdf(paste0(pdf_dir, "/LS8817vsBJ_Shared_NFKB.genes_vennDiagram.pdf"), width=4, height=15)
venn_data_shared <- list(LS8817_Shared=LS8817_NFkB$Shared$gene_id, BJ_Palbo=BJ_NFkB$`Palbo vs Cycling`, BJ_Doxo=BJ_NFkB$`Doxo vs Cycling`)
create_proportional_venn(sets=venn_data_shared, 
                         set_labels = c("LS8817 Shared", "BJ Palbo", "BJ Doxo"), 
                         colors = c("#E5C494",  "#ECCEE1" , "#D1DC8E") 
)
LS8817vsBJ_Shareddf <- get_intersections_named(venn_data_shared)
LS8817vsBJ_Shareddf$SASP <- ifelse(LS8817vsBJ_Shareddf$gene_id %in% row.names(BJ.anno_genes[BJ.anno_genes$SASP %in% "Yes", ]), "1", "0")

ggplot(LS8817vsBJ_Shareddf,
       aes(x = intersect_cat,
           y = gene_id,
           color = SASP,
           label = gene_id)) +
  geom_text(size = 3) +
  facet_wrap(~intersect_cat, scales="free")+
  scale_color_manual(values = c("0" = "black", "1" = "darkgreen")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(), 
    legend.position = "none"
  ) +
  labs(x = "", y = "")

venn_data_shared <- list(BJ_Palbo=BJ_NFkB$`Palbo vs Cycling`, BJ_Doxo=BJ_NFkB$`Doxo vs Cycling`)
create_proportional_venn(sets=venn_data_shared, 
                         set_labels = c("BJ Palbo", "BJ Doxo"), 
                         colors = c("#ECCEE1" , "#D1DC8E") 
)

BJ_Shareddf <- get_intersections_named(venn_data_shared)
BJ_Shareddf$SASP <- ifelse(BJ_Shareddf$gene_id %in% row.names(BJ.anno_genes[BJ.anno_genes$SASP %in% "Yes", ]), "1", "0")

ggplot(BJ_Shareddf,
       aes(x = intersect_cat,
           y = gene_id,
           color = SASP,
           label = gene_id)) +
  geom_text(size = 3) +
  facet_wrap(~intersect_cat, scales="free")+
  scale_color_manual(values = c("0" = "black", "1" = "darkgreen")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(), 
    legend.position = "none"
  ) +
  labs(x = "", y = "")
dev.off()
```
# make matrix of normalized counts across drug & time. 
```{r}
NormCountsMat <- counts(BJ_RNAdds, normalized=T)
# convert row names of matrix of normalized counts of off to on genes to gene symbol format. If there is no gene symbol format, keep ensembl format. 
row.names(NormCountsMat) <- gsub('\\.[0-9]*$', '', row.names(NormCountsMat))
row.names(NormCountsMat) <- ifelse(genes_anno_type[match(row.names(NormCountsMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(NormCountsMat), genes_anno_type[match(row.names(NormCountsMat), genes_anno_type$ensembl_id), ]$gene_id)
write.csv(NormCountsMat, file="/lustre/fs4/risc_lab/scratch/jyeung/Different_Cell_Lines_NFkB_Profiling_PDvsDoxo_Jan2026/post_STAR_alignment_analysis/BJ_HS_RNA_NormalizedCounts.csv")

# convert matrix of normalized counts of off to on genes into a dataframe with metaData information for ggplot to accept
NormCounts_df <- list()
for(i in 1:nrow(NormCountsMat)){
NormCounts_df[[i]] <- data.frame(Counts=NormCountsMat[i, ], sample=names((NormCountsMat[i, ])), Transcript= row.names(NormCountsMat)[i], metaData)
}
NormCounts_df <- do.call(rbind, NormCounts_df) # combine list into 1 giant dataframe
```

```{r}
NormCounts_df$Timepoint <- as.numeric(as.character(NormCounts_df$Timepoint))
# save plot to pdf
pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Different_Cell_Lines_NFkB_Profiling_PDvsDoxo_Jan2026/post_STAR_alignment_analysis/RNA_NormCountsMat_SASP.pdf", width=8, height=2)

ggplot(NormCounts_df[NormCounts_df$Transcript %in% row.names(anno_genes[anno_genes$NFKB_All %in% "Yes" & anno_genes$SASP %in% "Yes", ]) & !NormCounts_df$Treatment %in% "TNFa", ], aes(x=Timepoint, y=Counts, color=Treatment, shape=Biorep))+geom_point(size=3)+scale_color_manual(values=hm_cols_RNA$Treatment)+
facet_wrap(~Transcript, scales="free")+
ylab("Normalized Counts")+xlab("Length of Treatment (days)")+
  ggtitle("")+
  scale_y_continuous(n.breaks=5)+scale_x_continuous(n.breaks=5)+
  theme_classic()+theme(
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10), 
      axis.title.x = element_text(size = 10)
)

ggplot(NormCounts_df[NormCounts_df$Transcript %in% c("IL1A", "IL1B", "IL6"), ], aes(x=Timepoint, y=Counts, color=Treatment, shape=Biorep))+geom_point(size=3)+scale_color_manual(values=hm_cols_RNA$Treatment)+
facet_wrap(~Transcript, scales="free")+
ylab("Normalized Counts")+xlab("Length of Treatment (days)")+
  ggtitle("")+
  scale_y_continuous(n.breaks=5)+scale_x_continuous(n.breaks=5)+
  theme_classic()+theme(
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10), 
      axis.title.x = element_text(size = 10)
)

dev.off()
```

# RNA-seq: GSEA Analysis 
```{r RNAseq GSEA analysis}
ranked_res_forGSEA <- list()

for(i in 1:length(LFCResults_BJ_RNA)){
  
  rank <- LFCResults_BJ_RNA[[i]]
  
  # Fix NA p-values
  rank$pvalue[is.na(rank$pvalue)] <- 1
  
  # Avoid Inf
  rank$pvalue[rank$pvalue == 0] <- 1e-300
  
  # Remove NA logFC
  rank <- rank[!is.na(rank$log2FoldChange), ]
  
  # Compute ranking metric
  rank$rank_metric <- sign(rank$log2FoldChange) * -log10(rank$pvalue)
  
  # Remove non-finite values
  rank <- rank[is.finite(rank$rank_metric), ]
  
  # Sort
  rank <- rank[order(rank$rank_metric, decreasing = TRUE), ]
  
  # Create named vector
  ranked_res_forGSEA[[i]] <- rank$rank_metric
  names(ranked_res_forGSEA[[i]]) <- rownames(rank)
  
  # Convert to Entrez IDs
  entrez_ids <- make.unique(
    genes_anno_type[match(names(ranked_res_forGSEA[[i]]),
                          genes_anno_type$ensembl_id), ]$entrezid
  )
  
  names(ranked_res_forGSEA[[i]]) <- ifelse(
    is.na(entrez_ids),
    names(ranked_res_forGSEA[[i]]),
    entrez_ids
  )
  
  # 🔥 Remove NA names (GSEA will crash otherwise)
  valid <- !is.na(names(ranked_res_forGSEA[[i]]))
  ranked_res_forGSEA[[i]] <- ranked_res_forGSEA[[i]][valid]
}

# run GSEA analysis using: 
GSEA_Results_H <- list() # msigdb Hallmark database
for(i in 1:length(ranked_res_forGSEA)){
 GSEA_Results_H[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_H) 
}

names(GSEA_Results_H) <- names(LFCResults_BJ_RNA)
```

```{r}
# Get padj value of GSEA result for TNFA SIGNALING VIA NFKB pathway to add onto GSEA plots
GSEA_padj <- list(GSEA_Results_H$`Doxo vs Cycling`@result[GSEA_Results_H$`Doxo vs Cycling`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust, 
                  GSEA_Results_H$`Palbo vs Cycling`@result[GSEA_Results_H$`Palbo vs Cycling`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust, 
                  GSEA_Results_H$`TNFa vs Cycling`@result[GSEA_Results_H$`TNFa vs Cycling`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust)
                  
GSEA_padj <- lapply(GSEA_padj, function(x){format(round(x, 6), nsmall = 4)}) # round to 4 decimal places 

p1 <- gseaplot(GSEA_Results_H$`Doxo vs Cycling`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "Doxo vs Cycling: TNFA SIGNALING VIA NFKB", color.line="#D1DC8E", color="#D1DC8E", color.vline="#D1DC8E")+theme_classic()+ylim(-0.2, 0.8)+annotate("text", x = 30000, y = 0.8, label = paste0("padj=", GSEA_padj[[1]]))

p3 <- gseaplot(GSEA_Results_H$`Palbo vs Cycling`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "Palbo vs Cycling: TNFA SIGNALING VIA NFKB", color="#ECCEE1", color.line="#ECCEE1", color.vline="#ECCEE1")+theme_classic()+ylim(-0.2, 0.8)+annotate("text", x = 30000, y = 0.8, label = paste0("padj=", GSEA_padj[[2]]))

p5 <- gseaplot(GSEA_Results_H$`TNFa vs Cycling`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "+TNFa vs Untreated: TNFA SIGNALING VIA NFKB", color="darkblue", color.line="darkblue", color.vline="darkblue")+theme_classic()+ylim(-0.2, 0.95)+annotate("text", x = 30000, y = 0.8, label = paste0("padj=", GSEA_padj[[3]]))

pdf(paste0(pdf_dir, "/BJ_HS_GSEA_plots_Hallmark_TNFA.pdf"), width=5, height=4)
p1
p3
p5
dev.off()
```


